---
title: "PEM LiDAR Downloads"
author: "C Armour"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE, echo = FALSE}

library(sf)
require(tidyverse)
require(pbapply)
require(future)
require(parallel)
require(arcgisbinding)

list.files("functions", pattern = ".R$", full.names = T) %>% lapply(source)

```

First, set up your data path. The data path you set here will be used
throughout all other scripts. It is assumed that this path is different
to the automatic WD for the package.

```{r setup-folder-structure, include = FALSE}

data.path <- "D:/Kitimat_LiDAR/data/"

setup_folders(data.path = data.path)

```

At least once every six months, the index of available LiDAR files
should be re-downloaded from the portal. The code chunk below will check
the date of the most recent index and download a newer version if
applicable and an ArcGIS license is available. It returns the file path
of the relevant geopackage after this operation.

If you need to update manually for some reason, the instructions to do
this are as follows:

1.  Navigate to LidarBC Map Grid
    <https://governmentofbc.maps.arcgis.com/home/item.html?id=5f6a1f31212a4cb2826743d2e52ef02a>

2.  Select Open in ArcGIS Desktop -\> Open in ArcGIS Pro

3.   Double click item.pitemx file to open in ArcGIS

4.  Download point cloud index layer eg Point Cloud Index - 1:2,500 Grid

5.  Use "Export Features" tool and save to the "las_index_files" folder
    using the following format "las_index_YYYY-MM-DD.shp"

6.  Read this shapefile in as your index. Optional: export as gpkg.

You also need to read in the path to your AOI, which should be stored a
folder called "AOI" as a gpkg or shp.

```{r get-layers}

## Pull the AOI index
index <- make_gpkg_index() %>% st_read()

# Double check that the AOI is read in correctly as a single file
aoi <- sf::st_read("D:/Kitimat_LiDAR/AOI/Kitimat_PEM_AOI.shp")

```

Here, we query the lidar data. The data.path should already be specified
in the global environment but is named here as an argument for easier
troubleshooting. This function returns the index of tiles which are to
be downloaded.

```{r query-tiles, echo = FALSE}

aoi_index <- query_lidarbc(aoi = aoi, index = index, data.path = data.path)

```

Next is the download of the tiles. The following chunk takes the table
outputted from the "get-layers" chunk and parallelizes it for faster
download. You can set the number of cores in the "cores" variable.

```{r download}

cores = 6

options(timeout=600)

# Make a cluster with a number of cores, closing previous cores if needed
cl = makeCluster(cores)
# Load in required packages
clusterEvalQ(cl, {library(tidyverse); library(sf)})
# Export variables from global environment into each cluster
clusterExport(cl, c("aoi_index", "data.path"))

pblapply(
  aoi_index$task.no,
  cl = cl,
  FUN = function(j){
    
    tictoc::tic()
    # Get tile of interest
    tile <- filter(aoi_index, task.no == j)
    # Get URLs
    tile_url <- tile$s3Url
    rpt_url <- tile$acc_rpt_ur
    # Download las tile
    file_dest <- file.path(str_c(data.path,  'las/'), basename(tile_url))
    try(download.file(tile_url, file_dest, mode = 'wb'), silent = TRUE)
    # Download tile accuracy report
    if(is.character(rpt_url)){
      file_dest <- file.path(str_c(data.path,  'report/'), basename(rpt_url))
      try(download.file(rpt_url, file_dest, mode = 'wb'), silent = TRUE)
    }
    tictoc::toc()
    # Wait 30 seconds between downloads to avoid time out
    Sys.sleep(30)
    
  })

# Stop cluster
stopCluster(cl)

print('Downloads completed for area of interest')
```

This chunk will create batch files on your computer to pre-process the
data and then create CHMs and DEMs. It takes the previously specified
cores and data.path arguments. There is a resolution argument which you
can specify in the CHM and DEM functions - the default is set to 1m, 5m,
10m, and 20m for both.

```{r setup-batch-files}

setup_preprocess(data.path = data.path, cores = cores)

# chm.res = c(1, 3, 5, 7, 9)

setup_CHM(data.path = data.path, cores = cores)


# dem.res = c(1, 3, 5, 7, 9)

setup_DEM(data.path = data.path, cores = cores)

```
